### 2016-06-11 manuscript/AppendixS2.tex
Changes in code to fix pulling off wrong values of likelihood for AIC calcs in bearded seal example; changes to manuscript, and new Appendix S2 detailing trend surface model failure.  Paper version sent out for PRC (for the 2nd time).

% besdoc.tex V1.0, 17 March 2011

\documentclass[times,mee,doublespace,]{besauth2}
%%\documentclass[times,mee,]{besauth}

\newcommand{\journalnamelc}{British Ecological Society}
\newcommand{\journalabb}{British Ecological Society}
\newcommand{\journalclassshort}{BES}
%%\newcommand{\journalname}{British Ecological Society}
\usepackage{epstopdf,comment}

\usepackage{moreverb}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\usepackage{lineno}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\bibpunct[; ]{(}{)}{;}{}{}{,}

\def\volumeyear{2015}
\def\VOC{VO}

\begin{document}

\begin{center} \bf {\large Confronting preferential sampling in wildlife surveys: diagnosis and model-based triage}

\vspace{0.7cm}
Paul B. Conn$^{1,*}$, James T. Thorson$^2$, Devin S. Johnson$^1$
\end{center}
\vspace{0.5cm}

\rm
\small

$^1$ National Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA National Marine Fisheries Service, 7600 Sand Point Way NE, Seattle, WA 98115 USA; $^2$ Fisheries Resource Assessment and Monitoring Division (FRAM), Northwest Fisheries Science Center, National Marine Fisheries Service (NMFS),
NOAA, 2725 Montlake Boulevard E, Seattle, WA 98112, USA

$^*$paul.conn@noaa.gov

\large
\bigskip
\centerline{Appendix S2: Specific details on parameter estimation in bearded seal example}
\bigskip
\small

\linenumbers

\def\VAR{{\rm Var}\,}
\def\COV{{\rm Cov}\,}
\def\Prob{{\rm P}\,}

\begin{flushleft}

In addition to the models reported in the paper, we attempted to fit two different models with a trend surface specification for the preferential sampling process.  The models differed by whether or not predictive covariates were also modeled (the ice covariates described in the main article text), but otherwise the preferential sampling process was modeled similarly.  Specifically, when the probability $R_i$ that a location $i$ is selected for sampling is modeled as
\begin{eqnarray}
 \label{eq:R}
  R_i & \sim & \text{Bernoulli}(\textrm{expit}(\nu_i))),
\end{eqnarray}
where
\begin{eqnarray}
  \label{eq:nu}
  \nu_i & = & \beta_0^* + {\bf x}_i^* \boldsymbol{\beta}^*+\eta_i+{\bf B} \delta_i,
\label{eq:nu}
\end{eqnarray}
we implemented a trend surface model by writing $\bf{B}$ as
\begin{eqnarray}
  \label{eq:B}
  \textbf{B} & = & {\bf C}{\bf b}.
\end{eqnarray} 
Here, ${\bf b}$ is a column vector consisting of three estimated parameters, $b_0$, $b_1$, and $b_2$, and ${\bf C}$ is an $(S \times 3)$ design matrix constructed as follows:
\begin{eqnarray}
  \textbf{C} & = & \left[ 
    \begin{array}{ccc}
    1 & northing_1 & easting_1 \\
    1 & northing_2 & easting_2 \\
    \vdots & \vdots & \vdots \\
    1 & northing_S & easting_S 
    \end{array} 
  \right].
\end{eqnarray}
The northing and easting values are simply the $y$ and $x$ coordinates associated with each grid cell centroid on the particular geographical projection of the earth's surface that we were using (in our case, EASE Grid 2.0).

\hspace{0.5in} On the surface, there did not appear to be any numerical errors when fitting these trend surface models to bearded seal data.  The Hessian was invertible, and incorporation of the trend surface model resulted in a much lower AIC score than the other models (AIC scores were 19 and 26 points better than the highest ranked AIC model reported in the paper for the trend surface models omitting and including covariate effects, respectively).  However, when examining estimates graphically, it was clear that the trend surface models suffered from some parameter identification and/or aliasing issues (Fig. \ref{fig:bearded_RE_maps}).  Specifically, most variation in the probability of sampling was passed on from the $\boldsymbol{\eta}$ random effect to the shared random effect $\boldsymbol{\delta}$ such that it was difficult to disentangle the sampling process from the abundance process.  To our mind, this was a clear sign that the model was overparameterized, leading to poor estimation.

\raggedbottom

\bibliographystyle{bes}
\bibliography{master_bib}

\pagebreak

\begin{figure*} %[ht]
\begin{center}
\includegraphics[width=170mm]{bearded_RE_maps.pdf}
\caption{Estimates of spatial random effects for two models fitted to bearded seal count data.  Estimates from preferential sampling model $M_{cov=1,b=1}$ are provided in the first column of plots, while estimates from a second model with a trend surface specification for the preferential sampling parameter appear in the second column (this model also included covariate effects).  In the first column, the $\boldsymbol{\eta}$ random effects specific to the sampling probability model appear well estimated and to indicate a higher probability of sampling where transects were flown (as intended).  However, under the trend surface specification, these random effects are estimated to be close to zero, and the increased probability of sampling is passed on to the shared $\boldsymbol{\delta}$ random effects.} 
\label{fig:bearded_RE_maps}
\end{center}
\end{figure*}

\end{flushleft}
\end{document}

%http://www.plosone.org/article/info:doi/10.1371/journal.pone.0036527